### 2021-10-28 src-latex/parts/00-abstract.tex
Suggestion of structure

% !TeX root = ../main-paper.tex
The paper features a comparison of a selection (FIXME be more specific/enumerate) of named entity recognition (NER) techniques which can be reasonably used in the context of historical document analysis.
Such natural language processing techniques identify which class each text token belong to, like "person name", "location", "number" or "activity".
Introducing a new public dataset built from 19th century French directories, 
we perform a practical assessment of the accuracy of, and the annotation cost (considering various training set sizes) needed by each technique.
We also study the impact of real OCR noise for several OCR systems on the performance of NER.
Experiments suggest that FIXME APPROACH NAME is the most effective approach on the dataset we processed.
FIXME add numerical results
Results can be reproduced using data and code freely available at \url{https://github.com/soduco/paper-ner-bench-das22}
and the latest version of is paper can be obtained at \url{TODO arxiv URL}.