### 2022-01-17 src-latex/figs/eval-ocr-ner.tex
Add evaluation illutration

\documentclass[tikz,svgnames]{standalone}
\usepackage{newunicodechar}
\usepackage{amssymb,pifont}

\usetikzlibrary{matrix,positioning,calc}
\newunicodechar{⎵}{\verbvisiblespace}
\newunicodechar{❌}{{\color{red}\scriptsize\ding{53}}}

\begin{document}
\Pifont{pcr}

\begin{tikzpicture}
  \tikzset{alignment/.style={
      matrix of nodes, nodes={font=\tt, inner sep=0pt, minimum height=15pt}, column sep=0
    },
    taggedalignment/.style={alignment,
      column 1/.style={nodes={fill=red!20}},
      column 3/.style={nodes={fill=green!20}},
      column 5/.style={nodes={fill=blue!20}},
      column 7/.style={nodes={fill=yellow!20}},
    }
  }

  \node (A) { \verb|Alubert (Prosper),horlogerie,r, de la Ius-sienne,78| };

  \matrix[taggedalignment] (B) [right=3cm of A]
  {
    Alibert (Prosper)&,~ &horlogerie&,~ &r.~de la Jus- sienne&,~ &16&.\\
  };

  \matrix[alignment, row 2/.style={nodes={fill=none, gray!20}}] (C) [below=1.5cm of A]
  {
    Alibert (Prosper)&,~ &horlogerie&,~ &r.~de la Jus- sienne&,~ &16&.\\
    {||❌||||||||||||||}&{|❌}&{||||||||||}&{|❌}&{|❌❌||||||❌|||❌||||||}&{|❌}&{❌❌}&{❌}\\
    Alubert (Prosper)&,⎵ &horlogerie&,⎵  &r, de la Ius-⎵sienne&,7&8⎵ & ⎵\\
  };

  \matrix[taggedalignment, row 2/.style={nodes={fill=none, gray!20}}] (D) at ($ (C.east)!(B)!(C.west) $)
  {
    Alibert (Prosper)&,~ &horlogerie&,~ &r.~de la Jus- sienne&,~ &16&.\\
    {||❌||||||||||||||}&{|❌}&{||||||||||}&{|❌}&{|❌❌||||||❌|||❌||||||}&{|❌}&{❌❌}&{❌}\\
    Alubert (Prosper)&,⎵ &horlogerie&,⎵  &r, de la Ius-⎵sienne&,7&8⎵ & ⎵\\
  };

  \matrix[taggedalignment] (E1) [below=of D]
  {
    Alubert (Prosper)&,⎵ &horlogerie&,⎵  &r, de la Ius-⎵sienne&,7&8⎵ & ⎵\\
  };

  \matrix[taggedalignment] (E2) [below=of E1.west, anchor=west]
  {
    Alubert& (Prosper),⎵ &horlogerie&,⎵  &r, de la Ius-⎵sienne&,&78&⎵⎵\\
  };

  \node [above left=0cm of A, font=\it] {OCR Prediction};
  \node [above left=0cm of B, font=\it] {GOLD};

  \draw[-latex] (A.west) -- ++ (-1,0) |- (C-3-1) node[pos=0.25,left] { Alignment };

  \draw[-latex] (B) -- (D) node[midway,right] { Tag projection };
  \draw[-latex] (B.west) -- ++ (-1,0) |- (C-1-8);
  \draw[-latex] (C) -- (D);

  \draw[-latex] (C.south) -- ++(0,-1) node[midway,right,font=\small] {OCR Q.A.} node[below] {CER = 24\%};
  \draw[-latex] (D.south) -- ++(0,-1) node[midway,right,font=\small] {NER gt.};
  \draw[-latex] (C.south east) |- (E2) node[pos=0.75,below, font=\it] {NER Prediction};
  \draw[-latex] (E2.south) -- ++(0,-1) node[midway,right,font=\small] {NER Q.A.} node[below] {FIXME\%};
\end{tikzpicture}

\end{document}