### 2022-01-18 src-latex/main-paper.tex
WIP: section 4

\documentclass[runningheads,svgnames]{llncs}
%
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{cleveref}
% \usepackage{tikz}
% \usepackage{nth}
% \usepackage{subcaption}
\usepackage{booktabs}
% \usepackage{array}
\usepackage{orcidlink}
\usepackage{subcaption} % Addition
\usepackage{multirow} % Addition
\usepackage{tabularx} % Addition
\usepackage{booktabs} % Addition
\usepackage{scalerel}
\usepackage{cleveref}

% \usetikzlibrary{shadings,decorations,decorations.pathreplacing}
\renewcommand\UrlFont{\color{blue}\rmfamily}
% FIXME proper hyperref setup

\newcommand{\mcite}[1]{\text{\cite{#1}}}
\newcommand{\mref}[1]{\text{\ref{#1}}}

% \renewcommand{\baselinestretch}{0.97}

% Comment 
% FIXME : delete before submission :)
\newcommand{\bertrand}[1]{\textcolor{violet}{Bertrand:~#1}}
\newcommand{\nathalie}[1]{\textcolor{purple}{Nathalie:~#1}}
\newcommand{\edwin}[1]{\textcolor{pink}{Edwin:~#1}}

% Customize table rendering
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.4}

\begin{document}
%
\title{A Benchmark of Named Entity Recognition Approaches in Historical Documents\\
Application to 19$^{th}$ Century French Directories%
% FIXME "19$^{th}$-Century" with dash?
\thanks{This work was supported by the French National Research Agency:
Project SoDuCo, grant ANR-18-CE38-0013. ???FIXME need acknowledgments for directory sources???}}
%
\titlerunning{A Benchmark of NER Approaches in Historical Documents}
%
\author{LISTE PROVISOIRE PAR ORDRE ALPHA\\
%
Nathalie Abadie\inst{1}\orcidID{0000-0001-8741-2398} \and
Edwin Carlinet\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Joseph Chazalon\inst{3}\orcidID{2222--3333-4444-5555} \and
Bertrand Duménieu\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{LASTIG, Univ. Gustave Eiffel, IGN-ENSG, F-94160 Saint-Mandé, France\\
\email{nathalie-f.abadie@ign.fr}\\
\url{https://www.umr-lastig.fr/nathalie-abadie/} \and
INST, CITY, COUNTRY\\
\email{example@example.com}\\
\url{https://www.example.com} \and
CRH-EHESS, Paris, France\\
\email{bertrand.dumenieu@ehess.fr}\\ \url{http://crh.ehess.fr/index.php?5206}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{parts/00-abstract.tex}
\keywords{Historical documents \and Natural Language Processing \and Named Entity Recognition \and OCR noise \and Annotation cost.}
\end{abstract}

% ---- Sections ----
\input{parts/10-introduction.tex}
\input{parts/20-related-work.tex}
\input{parts/30-dataset.tex}
\input{parts/40-ner-approaches.tex}
\input{parts/50-xp-setup.tex}
\input{parts/60-results.tex}
\input{parts/80-conclusion.tex}

\section*{Acknowledments and Thanks}
This work is supported by the French National Research Agency (ANR), as part of the SODUCO project, under Grant ANR-18-CE38-0013.

The authors want to thank \textcolor{green}{Laureencia for annotations, Guillaume Thomas for data processing? Stéphane Bacciochi and Pascal Cristofoli for helping to create the gold reference dataset ?}

% ---- Bibliography ----
\bibliographystyle{splncs04}
\bibliography{ref}
\end{document}