### 2022-01-20 src-latex/parts/30-dataset.tex

% !TeX root = ../main-paper.tex
\section{Dataset}
\label{sec:dataset}

% Conventions nommage
% Dataset annoté à la main = "ground truth"
% ground truth aligné = "ner-reference"
% entrées pero aligné = "ner-pero"
% entrées tesseract alignées = "ner-tesseract"

%In this work, we  aim  at  producing  structured  spatio-temporal  data  from the information contained in the XIX\textsuperscript{th} century Parisian trade directories.
%To do this, it is therefore necessary to extract the text from the scans of the directories to be processed and to identify the entries in the various indexing lists used.
%Then, in each of these entries, the named entities they contain have to be extracted to produce structured spatio-temporal data representing the evolution over time of the people and businesses listed in these directories, their descriptions and their locations.
%Both of these extraction tasks require data to train and evaluate the OCR and NER approaches identified as potentially relevant: Tesseract and PERO OCR for the OCR task and Spacy and CamemBERT - pre-trained or fine-tuned - neural models for the NER task.

\subsection{A selection of Paris trade directories from 1798 to 1854}

The directories to be processed are stored in different libraries in Paris. They have been scanned independently and with various levels of quality. Moreover, they cover a wide period and have been produced by different publishers and printers. Therefore, their contents, indexes, layouts, methods of printing, etc. may vary a lot from one directory to another (see Figure \ref{fig:directories}). For training and testing purposes, our groundtruth dataset must be as representative as possible of the diversity of the available directories.

\begin{figure}[htb!]
	   \center{\includegraphics[width=0.9\textwidth]
	       {./images/DirectoryExcerpts2.png}}
	  \caption{\label{fig:directories} Examples of directory layouts and contents: 1) Duverneuil et La Tynna 1806 - index by name; 2) Deflandre 1828 - index by activity ; 3) Bottin 1851 - index by street name}
\end{figure}

For example, the Didot directory of 1854 is organised according to three indexes: by name, by activity, and by street name. The structure of the following entry is characteristic of its index by name: "Batton (D.-A.) \scalerel*{\includegraphics{./images/LH.png}}{H}, professeur au Conservatoire de musique et de déclamation, Saint-Georges, 47." This entry begins with a person's name, followed by the initials of their first name in parentheses.The symbol indicates that this person was awarded the Légion d'Honneur. The following part of this entry represents the person's activity,i.e., their profession, or social status, here "professor at the Conservatory of music and declamation". The street name and number where the person lives or carries out their activities are written at the end of the entry.

These pieces of information constituting the structure of the entries are the basic data that we want to extract, deduplicate, and structure to build a spatio-temporal database of XIX\textsuperscript{th} century Parisian inhabitants and businesses. Except for some potentially wordy activity descriptions, they correspond to named entities. However, while most entries contain the same types of named entities, their order and the way they are written vary from one directory or index to another. To provide examples of each entry structure, pages from each type have thus been annotated. %\footnote{In this work, the street name index entries have not been included as they are very different from the others. The strategy to be adopted to process them without loss of quality for the NER model will be the subject of future work.}.

\subsection{A dataset for OCR and NER evaluation}

\begin{figure}[tb]
    \includegraphics[width=\linewidth]{figs/protocol.pdf}
    \caption{Data extraction pipeline with two quality control checkpoints. }
    \label{fig.pipeline}
\end{figure}

%\joseph{TODO: separate entry detection, OCR and NER stages explicitely}

The simplified data extraction pipeline depicted in~\cref{fig.pipeline} processes the documents presented
in~\cref{sec:dataset}. First, the page layout extraction and entry segmentation are performed with a
semi-automated system and checked by a human. The resulting images, representing 8765 entries, are the inputs of a customizable two-stage pipeline.

%1. Entrée/Sortie de l'OCR
\textbf{OCR stage.} The first stage of the pipeline aims at extracting the raw text from the images. An OCR system runs
on the thumbnails of each segmented entry to extract its text. An entry might span over multiple text lines but is
always a single block. Thus, the most adapted mode is chosen when the OCR system allows for the detection mode (e.g. the
\emph{block} mode for \emph{Tesseract}). The glyphs used in this dataset might be unknown to the OCR. Some of them, like
\scalerel*{\includegraphics{./images/LH.png}}{H}, do not even have a unicode codepoint. Text normalization is thus required
on the OCR outputs and consists in projecting Unicode characters onto the latin-1 \emph{charset} (whenever it makes
sense), homogenize extra symbols (star and hand symbols) and alleviate some ambiguities (e.g. \emph{no} and \emph{n°}).

At the end of this first stage, an OCR Quality Assessment (Q.A.) between the \emph{normalized} groundtruth text and the
OCR output is performed.

%\joseph{[DONE] a word about the charset, and existence of glyphs unknown to the OCR}
%\joseph{[DONE] parler de l'étage d'extraction d'entrées qui est validé manuellement, et des quelques règles de simplification du charset pour homogénéiser entre les annotateurs / pallier des ambiguïtés dans les règles d'annotation (qu'on pourrait encore améliorer)}

%2. Entrée/Sortie
\textbf{NER Stage.} The second stage of the pipeline aims an extracting the named entities from a text with a NER
system. This text originates from the OCR outputs in a real-word scenario but might be groundtruth text in order to
evaluate the NER performance standalone. There are 5 type of entity to detect~(see.~\cref{tab.entities}).
The NER system has to classify non-overlapping parts of the text into one of these entity (or none of them). A second
Q.A. (namely NER Q.A.) is performed at the end of this stage between the groundtruth and the NER output.

Note that the dataset used to assess the NER stage is subset of the entries. Indeed, we need to ensure that the datasets
contain the same entries whichever the OCR used in the previous stage. Entries where the OCR produced an empty string
and those for which no entity could be projected from the groundtruth have to be ignored. We filter the set of entries by
keeping only the entries that are always valid at the end of the stage 1. At the cost of losing  5\% of the references
entries, we thus have 8341 valid entries at the input of the stage 2.

\begin{table}
\small
\begin{tabular}{llp{8cm}}
Entity & Count & Description\\    
PER & 8788 & a person name or a business name, usually referred to as several person names. First names, initials, or civility mentions are included. E.g.: "Alibert (Prosper)", "Allamand frères et Hersent", "Heurtemotte (Vve)", etc. \\
TITLE & 483 & an honorary title, either text, glyphs or a combination of the two. E.g: "O. \scalerel*{\includegraphics{./images/LH.png}} {H}" for "Officier de la Légion d'Honneur" or "\scalerel*{\includegraphics{./images/GMLondon.png}}{H}" to represent the Great Medal at the London exhibition.\\
ACT & 6472 & the profession or social status of a person summarized in the single concept of activity. E.g.: "horlogerie", "export en tous genres", "Conseiller d'Etat", "propriétaire".\\
LOC & 9709 & mostly street names ("r. de la Jussienne"), but may also be neighborhoods ("Marais du Temple") or any indirect spatial references ("au Palais Royal").\\
CARDINAL & 8747 & a street number, as part of an address ("16", "5 bis"), or a range of numbers (e.g. "23-25" or "5 à 9").\\
FT & 43 & a geographic entity type, used to give more details on a location, e.g. "boutique", "atelier", "fab." or "dépôt". \\
\end{tabular}
\caption{Entities to recognize in the dataset.}
\label{tab.entities}
\end{table}

% It can therefore be used to train and improve the entry detection, OCR and NER tools.
% book, page, box, text_ocr, ner_xml (+uid within a page)

%\joseph{TODO intégrer dans les sections avant ?}
%
%
% valid ner (ref \& pero \& tess): 8341
%
%TODO average entry len (in char)
%TODO average number of tags/entities per entry
%TODO lists actual dataset content: set of entries. For each entry:
%\begin{itemize*}
%    \item ref to original book, page, region
%    \item human transcription (OCR GT)
%    \item human NER (NER GT)
%    \item [GENERATED] OCR produced by Pero OCR, Tesseract v4, Kraken (precise model used)
%    \item [GENERATED] NER GT aligned on text predictions from Pero OCR, Tesseract v4, Kraken
%\end{itemize*}

%\joseph{TODO expliquer qu'il faut faire la fusion des sous-datasets NER}

%\subsection{Focus on how NER GT was aligned on noisy OCR}

\subsection{Metrics for OCR and NER Quality Assessment}

\begin{figure}[tb]
    \includegraphics[width=\linewidth]{figs/eval-ocr-ner.pdf}
    \caption{OCR and NER evaluation protocol example.}
    \label{fig.eval-ocr-ner}
\end{figure}

\textbf{OCR Q.A.}  The predicted text by the OCR system is aligned with the groundtruth text using standard tools from
Stephen V. Rice's thesis~\cite{santos.2019.wcmel,neudecker.2021.whdip}. The Character Error Rate (CER) is computed at
the entry level and at the global level, defined as the ratio between the number of errors
(insertions/deletions/substitutions) over the reference text length.

%In this benchmark, we consider 3 OCR systems well known for historical document analysis: Pero OCR, Tesseract, and Kraken.

\textbf{NER Q.A.}. The NER system outputs a text with tags that enclose the entities. To assess the quality of the
entity extraction, we rely on a technique similar as for the OCR evaluation to build the \emph{NER-target}. The
\emph{NER-target} is different from the groundtruth because it should not involve the errors committed during the
previous stages. The OCR text is first aligned with the groundtruth text to form the NER-\emph{input} (where
\emph{input} is a placeholder for \emph{pero} if the input text is from PERO, NER-\emph{tesseract},
NER-\emph{reference}\ldots). The tags of the groundtruth are then projected in the alignment on NER-\emph{input} to
provide the \emph{NER-target}. The NER system then runs on NER-\emph{input} and outputs the \emph{NER-prediction}. The
precision, recall, and f-measure (or f-score) are computed considering only the exact matches between entities of the
\emph{NER-target} and those from the \emph{NER-prediction}. Precision and recall are defined by \cref{eqn.prec-recall};
the f-measure is their harmonic mean.

\begin{align}
    \mathrm{precision} &= \frac{\text{\#exact matches}}{\text{\#entries in prediction}} & \mathrm{recall} &= \frac{\text{\#exact matches}}{\text{\#entries in reference}}
    \label{eqn.prec-recall}
\end{align}

The evaluation metrics are illustrated on~\cref{fig.eval-ocr-ner}. The OCR text and the groundtruth are first aligned to
evaluate the OCR accuracy. As there are 11 mismatches over 56 aligned characters, the CER is thus 24\%. This alignment
is then used to back-project the groundtruth tags to build the tagged NER-\emph{target}. Finally, the NER system runs on the
OCR text; its output is compared to the NER groundtruth. There is only 2 over 3 tags matching in the prediction (precision),
while only 2 over 4 entities are matched in the reference (recall). It follows an overall f-score of 0.4.