### 2022-02-28 rebuttal/rebuttal_response_max_500_words.txt
500 words exactly

We thank the reviewers for their comments. We will do our best to update the final version accordingly.

## Metrics and evaluation protocol

> [Reviewer 1] ICDAR 2011 Contest related to that topic

Thanks for pointing this out. They are essentially using a cross evaluation of all methods to perform end-to-end evaluation, just like us, but do not report NER evaluation, unlike us.

> [Reviewer 1] Qualitative analysis of influence of OCR quality on NER

Effects of the OCR noise are indeed non-uniform on NER results. The impact on structuring elements of the entries like punctuation symbols appears to be most significant, as they mark the separation between semantically homogeneous blocks. This is under further study.

> [Reviewer 1] how useful [projecting expected outcomes onto noisy OCR results] is 
> and whether [this] is not introducing biases in various ways, overestimating or underestimating 
> the effects of the NER labeling

We used standard metrics from previous works. Projecting the ground-truth labels on a noisy text was a sensible way to assess whether NERs could help to correct OCR errors. In our experiments, alignments usually make sense or can be detected as degenerated cases. However, it is very true that some borderline cases may cause trouble. We will be more explicit about these limitations.

> [Reviewer 2] Comparing two runs of PERO OCR together, or two runs of Tesseract together

Very good point, but we could not compete with the large, private training datasets used to produce the models we used. We chose to report, as objectively as possible, the performance of actual, off-the-shelf systems.

> [Reviewer 2] Reason for not reporting Kraken's performance on downstream NER

We chose not to report Kraken's performance as it would have drastically reduced the size of the labelled sets used to compare the performance of NER systems, because poor Kraken's performance hinders the projection of labels from the ground truth and creates degenerated cases which are skipped.

> [Reviewer 2] Experiment with artificial OCR noise in future work

We plan to perform such experiments, but as it is hard to generate realistic noise, we decided to first focus on real cases with no further assumptions.

## Paper organization

> [Reviewer 3] Section 2.3 should be a separate section about the pipeline

We have tried to extract this content as a new section as suggested, but this unbalanced sections, disrupted the explication flow, and breaks the 15-pages limits. As a consequence, the overall organization of the paper was left unchanged.

## Typos, inconsistencies and lack of clarity

> [Reviewer 1] Missing Figure 1 reference
> [Reviewer 1] you sometimes write "fig. n", "figure n", and "Figure n"
> [Reviewer 1] Tesseract spelling

Thanks, we fixed this.

> [Reviewer 3] What are "thumbnails" in page 7

We refer to "thumbnails" as image crops, they have the same resolution as the original image.

> [Reviewer 3] Description of NER QA is confusing
> [Reviewer 3] Figure 4 needs clarification

We clarified the relevant section and fig.4's caption.